---
title: "Hydraulics"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
toc_float: true
toc_depth: 3
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
---

```{r include=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
source("QUADP.R")
```

# Essential functions

## Conductivity as a function of water potential

The conductivity of an infinitesimally short xylem element can be described as a function of the water potential ($\Psi$), declining to half at a level $\Psi_{50}$.  
$$
k(\Psi) = 0.5 ^ {\left( \Psi / \Psi_{50} \right) ^ b}
$$
Plotted with different shape parameter values $b$.
```{r}
par_conductivity_std <- list(psi50 = -2, b = 2)

calc_conductivity <- function(psi, par){
  0.5^((psi/par$psi50)^par$b)
}

ggplot( data = tibble(x = 0), aes(x = x) ) +
  stat_function(fun = calc_conductivity, args = list(par = list(psi50 = -1, b = 1)), col = 'black') +
  stat_function(fun = calc_conductivity, args = list(par = list(psi50 = -1, b = 2)), col = 'green') +
  stat_function(fun = calc_conductivity, args = list(par = list(psi50 = -1, b = 5)), col = 'red') +
  xlim( -3, 0 ) +
  labs( title = "Conductivity as a function of water potential" )
```

## Total conductance as a function of leaf water potential

The total conductance along the soil-to-leaf pathway is then given by the integral 
$$
\int_{\Psi_S}^{\Psi_L}k(\Psi) d \Psi
$$

Plotted here for different soil water potentials.
```{r}
calc_conductance = function(dpsi, psi_soil, ...){
  -integrate(calc_conductivity, lower = psi_soil, upper = (psi_soil - dpsi), ...)$value
}

# integral_P = function(dpsi, psi_soil, psi50, b, ...){
#   -integrate(P, psi50=psi50, b=b, lower = psi_soil, upper = (psi_soil - dpsi), ...)$value
# }

tibble( dpsi = seq(0, 5, length.out = 30)) %>% 
  mutate(total_conductance_soil1 = purrr::map_dbl(dpsi, ~calc_conductance(., psi_soil = -1, par = par_conductivity_std))) %>% 
  mutate(total_conductance_soil2 = purrr::map_dbl(dpsi, ~calc_conductance(., psi_soil = -2, par = par_conductivity_std))) %>% 
  ggplot() +
  geom_line(aes( x = dpsi, y = total_conductance_soil1)) +
  geom_line(aes( x = dpsi, y = total_conductance_soil2), col = 'red') +
  labs( title = "Total conductance as a function of leaf water potential" )
```

Specify plant hydraulic parameters that are held constant at the subdaily to annual time scale.
```{r}
par_plant_std = list(
    Ks0=1e-12, 
    v_huber=1e-4, 
    height=10, 
    conductivity_scalar=1,  # Scales (Ks0*v_huber/H)
    psi50 = -2, 
    b=2
    )
```

With these parameters, typical $E$ (per unit leaf area) would be around:
```{r}
E = par_plant_std$Ks0 * par_plant_std$v_huber / par_plant_std$height / par_env_std$viscosity_water * 1e6

m_per_s_to_mm_per_day = 8.64e7

cat("E = ", E, " m^3/m2/s", "\n",
"E = ", E*m_per_s_to_mm_per_day, " mm/day", "\n",
"E = ", E*m_per_s_to_mm_per_day*365, " mm/year", sep = "")
```

Thus, a forest with $LAI = 4$ will transpire 
```{r}
cat("E_forest = ", E*m_per_s_to_mm_per_day*365*4 ," mm/year")
```

The dependence of $E$ on soil water potential and the soil-to-leaf water potential difference $\Delta \Psi$ is given by:
```{r}
# calc_transpiration <- function(dpsi, psi_soil, par_plant, par_env, ...){
#   par_plant$conductivity_scalar * par_plant$Ks0 * par_plant$v_huber / par_plant$height / par_env$viscosity_water * 1e6 * integral_P(dpsi, psi_soil, par_plant$psi50, par_plant$b, ...)
# }
calc_transpiration <- function(dpsi, psi_soil, par_plant, par_env, ...){
  par_plant$conductivity_scalar * par_plant$Ks0 * par_plant$v_huber / par_plant$height / par_env$viscosity_water * 1e6 * calc_conductance(dpsi, psi_soil, par = par_plant)
}

list( dpsi = seq(0, 5, length.out = 30), psi_soil = c(0,-1,-2)) %>% 
  cross_df() %>% 
  mutate(E = purrr::pmap_dbl(data.frame(dpsi=dpsi, psi_soil=psi_soil), calc_transpiration, par_plant=par_plant_std, par_env=par_env_std)) %>%
  ggplot() +
  geom_line(aes( x = dpsi, y =  E*m_per_s_to_mm_per_day, group=psi_soil, colour=as.factor(psi_soil)), size=1) +
  scale_color_manual(values = c("red", "yellow3", "green3")) +
  ylab("E (mm/day)")
```

# Short-term

## Short-term without cavitation cost

Short term regulation of $g_s$ at time scales of minutes to days.

- **Given**: $V_{\mathrm{cmax}}, \nu_H, k_{s0}, H, \rho_s, \Delta \Psi^\ast$
- **Environment**: $\Psi_s, \eta, D$
- **Principle**: Maintain a regulated $g_s^\ast$ so that $\Delta \Psi(g_s) = \Delta \Psi^\ast$.

This yields $g_s^\ast$ as a function of hydraulic parameters, the "isohydricity" constant $\Delta \Psi^\ast$, and the environment:
$$
g_s^\ast = \frac{\nu_H \; k_{s0}}{1.6 H \eta D} \int_{\Psi_s}^{\Psi_s - \Delta\Psi^\ast} k(\Psi) \; d\Psi
$$

```{r}
calc_gs_star <- function(psi_soil, vpd, viscosity, dpsi_star, v_huber, par, par_conductivity){
  (v_huber * par$conductivity_base / (1.6 * par$height * viscosity * vpd)) * calc_conductance(dpsi = dpsi_star, psi_soil = psi_soil, par = par_conductivity_std) 
}
# par_conductance_std <- list(conductivity_base = 1, height = 1 )
```

### Response to soil drying

The regulated stomatal conductance $g_s^\ast$ responds to a drying soil as shown below (plotted as a function of soil water potential).
```{r}
tibble(psi_soil = seq(0, -5, length.out = 30)) %>% 
  mutate(gs_star_1 = purrr::map_dbl(psi_soil, ~calc_gs_star(., vpd = 100, viscosity = 1, dpsi_star = 1, v_huber = 1, par = par_conductance_std, par_conductivity = par_conductivity_std))) %>% 
  mutate(gs_star_2 = purrr::map_dbl(psi_soil, ~calc_gs_star(., vpd = 100, viscosity = 1, dpsi_star = 2, v_huber = 1, par = par_conductance_std, par_conductivity = par_conductivity_std))) %>% 
  ggplot() + 
  geom_line(aes(x = psi_soil, y = gs_star_1)) +
  geom_line(aes(x = psi_soil, y = gs_star_2, col = 'red')) +
  labs(title = "Regulated stomatal conductance", subtitle = "At two different dpsi_star levels")
```

With soil water potential, expressed as a function of volumetric soil water content, using (from BiomeE):
$$
\Psi_s = \Psi_\text{sat} \left( \frac{W_\text{FC}}{W}\right) ^b
$$
```{r}
## functional form and values from BiomeE model
calc_psi_soil <- function(w_vol, par){
  par$psi_soil_sat * (par$w_vol_fc / w_vol)^par$b
}
par_psi_soil_std = list(psi_soil_sat = -0.6, w_vol_fc = 0.4, b = 1)  # b is 2.2 in BiomeE
```

... it can be plotted as a function of soil volumetric water content.
```{r}
df_w_vol <- tibble( w_vol = seq(0.05, 0.6, length.out = 100)) %>% 
  mutate(psi_soil = calc_psi_soil(w_vol, par_psi_soil_std)) %>% 
  mutate(gs_star_1 = purrr::map_dbl(psi_soil, ~calc_gs_star(., vpd = 100, viscosity = 1, dpsi_star = 1, v_huber = 1, par = par_conductance_std, par_conductivity = par_conductivity_std)))

df_w_vol %>% 
  ggplot() + 
  geom_line(aes(x = w_vol, y = gs_star_1)) +
  labs(title = "Regulated stomatal conductance", subtitle = "As a funtion of volumetric soil water content")
```

This yields a GPP response to drying soil.
```{r}
calc_assim <- function(gs, vcmax, ca, par){
  
  ## Rubisco is limiting
  ## Solve Eq. system
  ## A = gs (ca- ci)
  ## A = Vcmax * (ci - gammastar)/(ci + Kmm)
  
  ## This leads to a quadratic equation:
  ## A * ci^2 + B * ci + C  = 0
  ## 0 = a + b*x + c*x^2
  
  ## with
  A <- -1.0 * gs
  B <- gs * ca - gs * par$kmm - vcmax
  C <- gs * ca * par$kmm + vcmax * par$gammastar
  
  ci <- QUADM(A, B, C)
  a_c <- vcmax * (ci - par$gammastar) / (ci + par$kmm)
  
  return(a_c)
}

## Set P-model parameters
beta <- 146          # unit cost ratio a/b
c_cost <- 0.41
gamma <- 0.105       # unit cost ratio c/b
kphio <- 0.05        # quantum yield efficiency
c_molmass <- 12.0107 # molar mass, g / mol

## Define environmental conditions
tc <- 20             # temperature, deg C
ppfd <- 300          # mol/m2/d
vpd  <- 1000         # Pa
co2  <- 400          # ppm
elv  <- 0            # m.a.s.l.
fapar <- 1           # fraction

out_analytical <- rpmodel::rpmodel(
  tc             = tc,
  vpd            = vpd,
  co2            = co2,
  elv            = elv,
  kphio          = kphio,
  beta           = beta,
  fapar          = fapar,
  ppfd           = ppfd,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE
  )

par_photosynth_std <- list(
  kmm = out_analytical$kmm,
  gammastar = out_analytical$gammastar
  )
```

```{r}
df_w_vol <- df_w_vol %>% 
  mutate(a_c = purrr::map_dbl(gs_star_1, ~calc_assim(., vcmax = out_analytical$vcmax, ca = out_analytical$ca, par = par_photosynth_std)))

df_w_vol %>% 
  ggplot() + 
  geom_line(aes(x = w_vol, y = a_c)) +
  labs(title = "Assimilation", subtitle = "As a funtion of volumetric soil water content")
```

### Response to VPD

```{r}
df_vpd <- tibble( vpd = seq(0, 1000, length.out = 100)) %>% 
  mutate(gs_star = purrr::map_dbl(vpd, ~calc_gs_star(-1, vpd = ., viscosity = 1, dpsi_star = 1, v_huber = 1, par = par_conductance_std, par_conductivity = par_conductivity_std))) %>% 
  mutate(a_c_inst = purrr::map_dbl(gs_star, ~calc_assim(., vcmax = out_analytical$vcmax, ca = out_analytical$ca, par = par_photosynth_std))) 

df_vpd %>% 
  ggplot() + 
  geom_line(aes(x = vpd, y = a_c_inst)) +
  labs(title = "Assimilation A", subtitle = "As a funtion of VPD")
```

## Short-term with cavitation cost

xxx

# Mid-term

On time scales of weeks to months, $V_{\mathrm{cmax}}$ and $\nu_H$ are acclimated/regulated to satisfy some optimality criterion. 

At this time scale, other hydraulic parameters are regarded as constant, while stomatal conductance is instantaneously regulated with fast-varying environmental conditions ($g_s^\ast$).

- **Given**: $k_{s0}, H, \rho_s, A_s, \Delta \Psi^\ast$
- **Environment**: $\Psi_s, \eta, D$
- **Principle**: Optimise $(V_{\mathrm{cmax}}, \nu_H)$ with $g_s^\ast = f(\Delta \Psi^\ast, \nu_H)$

A least-cost (LC) principle could be formulated as
$$
 b V_{\mathrm{cmax}} / A + (cA_s /\nu_H)/A = min.
$$
The second term is the leaf construction cost which is a function of the Huber value $\nu_H$ and the sapwood cross-sectional area $A_s$. Note that the water cost is implicitly included in $A$ because it's a function of the regulated $g_s^\ast$.

### Simplification

Treat $\nu_H$ as a constant and optimise $V_{\mathrm{cmax}}$ so that net assimilation is maximised:
$$
 A_C - b V_{\mathrm{cmax}} = max.
$$
Does it even have a maximum?
```{r}
fn_target <- function(vcmax, gs_star, ca, par_cost, par_photosynth, do_optim = FALSE){
  ## Required parameter sets:
  ## par_cost        : b
  ## par_photosynth  : kmm, gammastar
  
  ## Rubisco-limited assimilation
  a_c <- calc_assim(gs_star, vcmax, ca, par = par_photosynth)
  
  ## Profit Maximisation
  out <- a_c - par_cost$b * vcmax
  
  if (do_optim){
    return(-out)
  } else {
    return(out)
  }
}  

optim_vcmax_midterm_constvhuber <- function(fn_target, gs_star, ca, par_cost, par_photosynth, vcmax_init, return_all = FALSE){

  out_optim <- optimr::optimr(
    par       = vcmax_init,
    lower     = 0.00001 * vcmax_init,
    upper     = 100000 * vcmax_init,
    fn        = fn_target,
    gs_star   = gs_star, 
    ca        = ca,
    par_cost  = par_cost, 
    par_photosynth = par_photosynth,
    do_optim  = TRUE,
    method    = "L-BFGS-B",
    control   = list( maxit = 100, maximize = TRUE )
  )
  
  out_optim$value <- -out_optim$value
  
  if (return_all){
    out_optim
  } else {
    return(out_optim$par)
  }
}


df_test <- tibble( vcmax = seq(out_analytical$vcmax * 0.1, out_analytical$vcmax * 10, length.out = 30) ) %>% 
  mutate(assim = purrr::map_dbl(vcmax, ~calc_assim(gs = 1, vcmax = ., ca = out_analytical$ca, par = par_photosynth_std))) %>% 
  mutate(target = purrr::map_dbl(vcmax, ~fn_target(., gs_star = 1, ca = out_analytical$ca, par_cost  = list(b = 0.05), par = par_photosynth_std)))

out_midterm_simpl <- optim_vcmax_midterm_constvhuber(fn_target, gs_star = 1, ca = out_analytical$ca, par_cost  = list(b = 0.05), par_photosynth = par_photosynth_std, vcmax_init = out_analytical$vcmax, return_all = TRUE)

df_test %>% 
  ggplot(aes(x = vcmax, y = target)) +
  geom_line() +
  geom_point(aes(x = out_midterm_simpl$par, y = out_midterm_simpl$value), col = 'red')
```
Yes. Ok.

Btw: A least-cost criterion ($b V_{\mathrm{cmax}} / A_C = min.$) doesn't work here

This allows us to predict how $V_{\mathrm{cmax}}$ would acclimate to a drying soil.
```{r}
df_w_vol <- tibble( w_vol = seq(0.05, 0.6, length.out = 100)) %>% 
  mutate(psi_soil = calc_psi_soil(w_vol, par_psi_soil_std)) %>% 
  mutate(gs_star = purrr::map_dbl(psi_soil, ~calc_gs_star(., vpd = 100, viscosity = 1, dpsi_star = 1, v_huber = 1, par = par_conductance_std, par_conductivity = par_conductivity_std))) %>% 
  mutate(vcmax_opt = purrr::map_dbl(gs_star, ~optim_vcmax_midterm_constvhuber(fn_target, gs_star = ., ca = out_analytical$ca, par_cost  = list(b = 0.05), par_photosynth = par_photosynth_std, vcmax_init = out_analytical$vcmax))) %>% 
  mutate(a_c_accl = purrr::map2_dbl(gs_star, vcmax_opt, ~calc_assim(.x, vcmax = .y, ca = out_analytical$ca, par = par_photosynth_std))) %>%  
  mutate(a_c_inst = purrr::map_dbl(gs_star, ~calc_assim(., vcmax = 0.886016918, ca = out_analytical$ca, par = par_photosynth_std)))

df_w_vol %>% 
  ggplot() + 
  geom_line(aes(x = w_vol, y = vcmax_opt)) +
  labs(title = "Acclimated Vcmax", subtitle = "As a funtion of volumetric soil water content")

df_w_vol %>% 
  tidyr::gather("source", "a_c", c(a_c_accl, a_c_inst)) %>% 
  ggplot() + 
  geom_line(aes(x = w_vol, y = out_opt)) +
  labs(title = "Regulated Vcmax", subtitle = "As a funtion of volumetric soil water content")

df_w_vol %>% 
  ggplot() + 
  geom_line(aes(x = psi_soil, y = out_opt)) +
  labs(title = "Regulated Vcmax", subtitle = "As a funtion of soil water potential")
```


### Instantaneous vs. acclimated

The instantaneous dependence of $A$ on VPD goes with $1/D$. This is probably not reasonable. Compare this to the acclimated response.
```{r}
df_vpd <- tibble( vpd = seq(0, 1000, length.out = 100)) %>% 
  mutate(gs_star = purrr::map_dbl(vpd, ~calc_gs_star(-1, vpd = ., viscosity = 1, dpsi_star = 1, v_huber = 1, par = par_conductance_std, par_conductivity = par_conductivity_std))) %>% 
  mutate(a_c_inst = purrr::map_dbl(gs_star, ~calc_assim(., vcmax = out_analytical$vcmax, ca = out_analytical$ca, par = par_photosynth_std))) %>% 
  filter(!is.infinite(gs_star)) %>% 
  mutate(vcmax_opt = purrr::map_dbl(gs_star, ~optim_vcmax_midterm_constvhuber(fn_target, gs_star = ., ca = out_analytical$ca, par_cost  = list(b = 0.05), par_photosynth = par_photosynth_std, vcmax_init = out_analytical$vcmax))) %>% 
  mutate(a_c_accl = purrr::map2_dbl(gs_star, vcmax_opt, ~calc_assim(.x, vcmax = .y, ca = out_analytical$ca, par = par_photosynth_std)))

df_vpd %>% 
  tidyr::gather("source", "a_c", c(a_c_accl, a_c_inst)) %>% 
  ggplot() + 
  geom_line(aes(x = vpd, y = a_c, linetype = source)) +
  labs(title = "Assimilation  vs. instantaneous A", subtitle = "As a funtion of VPD")
```




XXX todo: 

- Plot instantaneous $A$ response to soil drying (constant Vcmax) and acclimated $A$ (with acclimated Vcmax).
- Is assimilation reduced in proportion to conductance?
- Plot acclimated assimilation as a function of VPD.
