---
title: "Compare P-model implementations"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig_width: 30 
fig_height: 20 
vignette: >
  %\VignetteIndexEntry{test_implementations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
# output:
#   html_document:
#     toc: true
#     toc_float: true
#     toc_depth: 3
# # output:
# #   pdf_document:
# #     toc: true
# #     toc_depth: 2
# header-includes:
#    - \usepackage{amsmath}
# # bibliography: bibliography.bib
---

## Environment

Load the `rsofun` package. This contains all the necessary wrapper functions to set up and run SOFUN and read its output. 
```{r message=FALSE, echo=FALSE, warning=FALSE}
library(rsofun)
library(rpmodel)

## load all rsofun dependencies
load_dependencies_rsofun()

## other crap
knitr::opts_knit$set( root.dir = rprojroot::find_rstudio_root_file() ) # does not work properly
# if (is.null(options()$rsofun.dir.sofun)) rlang::abort( "Option rsofun.dir.sofun not set. Do so by `options( list( rsofun.dir.sofun=string_path_where_sofun_is ))`" )
options( list( rsofun.dir.sofun="~/sofun/" ))
```

## Run implementations for a temperature range

Varying temperature from 0 to 35 degrees Celsius. All other factors are held constant with:

- `ppfd = 800`
- `vpd = 1000`
- `co2 = 400`
- `elv = 0`

The quantum yield efficiency is set to `kphio = 0.05` for the `"wanghan"` method, and 0.257 for the `"smith"` method.

```{r, eval=TRUE, message=FALSE, warning=FALSE, fig.width=12}
ppfd <- 800
vpd  <- 1000
co2  <- 400
elv  <- 0
fapar <- 1

## R, Vcmax based on Wang Han's formulation
pmodel_stdrd_R <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~rpmodel::rpmodel( tc = ., vpd = vpd, co2 = co2, elv = elv, kphio = 0.05, fapar = fapar, ppfd = ppfd, method_optci="prentice14", method_jmaxlim = "wang17", do_ftemp_kphio = FALSE ) 
  )

pmodel_smith <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~rpmodel::rpmodel( tc = ., vpd = vpd, co2 = co2, elv = elv, kphio = 0.257, fapar = fapar, ppfd = ppfd, method_optci="prentice14", method_jmaxlim = "smith19", do_ftemp_kphio = FALSE ) 
  )

## Fortran, Vcmax basedon Wang Han's formulation
## update quantum yield parameter in file
#params_opt <- readr::read_csv( paste0( path.package("rsofun"), "/extdata/params_opt_kphio_soilm_global.csv" ) )
params_opt <- tibble(kphio = 0.05)
nothing <- update_params( params_opt, options()$rsofun.dir.sofun )

pmodel_fortran <- purrr::map( as.list( seq( 0, 35, length.out = 100 ) ), 
  ~pmodel( temp = ., vpd = vpd, co2 = co2, ppfd = ppfd, fapar = fapar, elv = elv, implementation = "fortran", sofundir = options()$rsofun.dir.sofun ) )
```

## Evaluation by temperature

Below, I'm comparing different variables calculated in the different implementations, as a function of temperature (to which photosynthesis acclimated to). 'rsofun standard' refers to what is implemented in rsofun, based on Wang Han et al., 2017 and originally adopted from the GePiSaT code. 'Smith my implementation' is based on Beni's adoptation of Nick Smith's code within rsofun. The temperature-dependence of quantum yield efficiency is not accounted for in any of the calculations (argument `do_ftemp_kphio = FALSE` in `rpmode()` function calls).

### Michaelis-Menten K

```{r, eval=TRUE, message=FALSE, warning=FALSE}
kmm_stdrd_R   <- pmodel_stdrd_R   %>% purrr::map_dbl("kmm")
kmm_fortran   <- pmodel_fortran   %>% purrr::map_dbl("kmm")
kmm_smith     <- pmodel_smith     %>% purrr::map_dbl("kmm")

plot(  seq( 0, 35, length.out = 100 ), kmm_stdrd_R, type = "l", xlab = "Temperature (deg C)", ylab = "Michaelis-Menten K", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), kmm_fortran, col="green", lwd=3 )
lines( seq( 0, 35, length.out = 100 ), kmm_smith, col="blue", lwd=1 )
legend( "topright", c("rsofun standard (R)", "rsofun standard (Fortran)", "Smith my implementation"), lty = 1, col = c("black", "green", "blue" ), lwd=c(6,3,1), bty = "n")
```

### $c_i$

```{r, eval=TRUE, message=FALSE, warning=FALSE}
ci_stdrd_R   <- pmodel_stdrd_R   %>% purrr::map_dbl("ci")
ci_fortran   <- pmodel_fortran   %>% purrr::map_dbl("ci")
ci_smith     <- pmodel_smith     %>% purrr::map_dbl("ci")

plot(  seq( 0, 35, length.out = 100 ), ci_stdrd_R, type = "l", xlab = "Temperature (deg C)", ylab = "ci", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), ci_smith, col="green", lwd=3 )
lines( seq( 0, 35, length.out = 100 ), ci_smith, col="blue", lwd=1 )
legend( "topright", c("rsofun standard (R)", "rsofun standard (Fortran)", "Smith my implementation"), lty = 1, col = c("black", "green", "blue" ), lwd=c(6,3,1), bty = "n")
```

### $c_i:c_a$

```{r, eval=TRUE, message=FALSE, warning=FALSE}
chi_stdrd_R   <- pmodel_stdrd_R   %>% purrr::map_dbl("chi")
chi_fortran   <- pmodel_fortran   %>% purrr::map_dbl("chi")
chi_smith     <- pmodel_smith     %>% purrr::map_dbl("chi")

plot(  seq( 0, 35, length.out = 100 ), chi_stdrd_R, type = "l", xlab = "Temperature (deg C)", ylab = "ci:ca", lwd=6 )
lines( seq( 0, 35, length.out = 100 ), chi_fortran, col="green", lwd=3 )
lines( seq( 0, 35, length.out = 100 ), chi_smith, col="blue", lwd=1 )
legend( "topright", c("rsofun standard (R)", "rsofun standard (Fortran)", "Smith my implementation"), lty = 1, col = c("black", "green", "blue"), lwd=c(6,3,1), bty = "n")
```

### Vcmax

```{r, eval=TRUE, message=FALSE, warning=FALSE}
vcmax_stdrd_R   <- pmodel_stdrd_R   %>% purrr::map_dbl("vcmax")
vcmax_fortran   <- pmodel_fortran   %>% purrr::map_dbl("vcmax")
# vcmax_smith     <- pmodel_smith     %>% purrr::map_dbl("vcmax_star")

plot(  seq( 0, 35, length.out = 100 ), vcmax_stdrd_R, type = "l", xlab = "Temperature (deg C)", ylab = "Vcmax", lwd=6, ylim=c(0,max(vcmax_stdrd_R)) )
lines( seq( 0, 35, length.out = 100 ), vcmax_fortran, col="green", lwd=3 )
legend( "topleft", c("rsofun standard (R)", "rsofun standard (Fortran)"), lty = 1, col = c("black", "green"), lwd=c(6,3), bty = "n" )
# lines( seq( 0, 35, length.out = 100 ), vcmax_smith, col="blue", lwd=1 )
# legend( "topleft", c("rsofun standard (R)", "rsofun standard (Fortran)", "Smith my implementation"), lty = 1, col = c("black", "green", "blue"), lwd=c(6,3,1), bty = "n" )
```


### Light use efficiency

```{r, eval=TRUE, message=FALSE, warning=FALSE}
lue_stdrd_R   <- pmodel_stdrd_R   %>% purrr::map_dbl("lue")
lue_fortran   <- pmodel_fortran   %>% purrr::map_dbl("lue")
lue_smith     <- pmodel_smith     %>% purrr::map_dbl("lue")

plot(  seq( 0, 35, length.out = 100 ), lue_stdrd_R, type = "l", xlab = "Temperature (deg C)", ylab = "lue", lwd=6, ylim=c(0,max(lue_smith)) )
lines( seq( 0, 35, length.out = 100 ), lue_fortran, col="green", lwd=3 )
legend( "topright", c("rsofun standard (R)", "rsofun standard (Fortran)"), lty = 1, col = c("black", "green"), lwd=c(6,3), bty = "n")
# lines( seq( 0, 35, length.out = 100 ), lue_smith, col="blue", lwd=1 )
# legend( "topright", c("rsofun standard (R)", "rsofun standard (Fortran)", "Smith my implementation"), lty = 1, col = c("black", "green", "blue"), lwd=c(6,3,1), bty = "n")
```


