---
title: "Usage"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{Running the P-model in R}
---

The rpmodel package implements the P-model as described in Stocker et al. (2019) *Geosci. Mod. Dev.* The main function available through the package is `rpmodel()`, which returns a list of quantities (see `?rpmodel`) for a given set of inputs. An additional set of important functions that are used within `rpmodel()` are also available through this package. Usage examples are given below.

## Example P-model run

Let's run the P-model, without $J_{\text{max}}$ limitation (argument `method_jmaxlim = "none"`), for one point. The set of inputs, being temperature (`tc`), photosynthetic photon flux density (`ppfd`), vapour pressure deficit (`vpd`), ambient CO$_2$ (`co2`), elevation (`elv`), and fraction of absorbed photosynthetically active radiation (`fapar`). The quantum yield efficiency parameter is provided as an argument (`kphio`) and corresponds to $\widehat{\varphi_0}$ in Stocker et al. (2019) if the temperature-dependence of this parameter is ignored (argument `do_ftemp_kphio = FALSE`, corresponding to simulation setup 'ORG' in Stocker et al. (2019)), or to $\widehat{c_L}$ if the temperature-dependence of the quantum yield efficiency is included (argument `do_ftemp_kphio = TRUE`, used in simulation setups 'BRC' and 'FULL' in Stocker et al. (2019)). By default the optional argument `do_soilmstress` is set to `FALSE`, meaning that the empirical soil moisture stress function is not included. The unit cost ratio ($\beta$ in Stocker et al. (2019)) is given by argument `beta`.

To run the `rpmodel()` function we can do:
```{r}
library(rpmodel)
out_pmodel <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 300,          # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  kphio          = 0.05,         # quantum yield efficiency,
  beta           = 146,          # unit cost ratio a/b,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE,
  do_soilmstress = FALSE,
  verbose        = TRUE
  )
print(out_pmodel)
```

Here, we specified the model paramters (arguments `beta` and `kphio`). This overrides the defaults, where `rpmodel()` uses the parameters as calibrated by Stocker et al. (2019), depending on the choices for arguments `do_ftemp_kphio` and `do_soilmstress`:
```{r eval=FALSE}
kphio = ifelse(do_ftemp_kphio, ifelse(do_soilmstress, 0.087182, 0.081785), 0.049977)
beta = 146.0
apar_soilm = 0.0
bpar_soilm = 0.73300
```

The function returns a list of variables (see also man page by `?rpmodel`), including $V_{\mathrm{cmax}}$, $g_s$, and all the parameters of the photosynthesis model ($K$, $\Gamma^{\ast}$), which are all internally consistent, as can be verified for...
$$
c_i = c_a - A / g_s = \chi c_a
$$

```{r}
c_molmass <- 12.0107  # molecular mass of carbon
kphio <- 0.05         # quantum yield efficiency, value as used in the function call to rpmodel()
ppfd <- 300           # mol/m2/d, value as used in the function call to rpmodel()
fapar <- 1            # fraction, value as used in the function call to rpmodel()
print( out_pmodel$ci )
print( out_pmodel$ca - (out_pmodel$gpp / c_molmass) / out_pmodel$gs )
print( out_pmodel$ca * out_pmodel$chi )
```
Yes. 

And for...
$$
A = V_{\text{cmax}} \frac{c_i-\Gamma^{\ast}}{c_i + K} = \phi_0 I_{\text{abs}} \frac{c_i-\Gamma^{\ast}}{c_i + 2 \Gamma^{\ast}} = g_s (c_a - c_i)
$$

```{r}
print( out_pmodel$gpp / c_molmass )
print( out_pmodel$vcmax * (out_pmodel$ci - out_pmodel$gammastar) / (out_pmodel$ci + out_pmodel$kmm ))
print( out_pmodel$gs * (out_pmodel$ca - out_pmodel$ci) )

print( kphio * ppfd * fapar * (out_pmodel$ci - out_pmodel$gammastar) / (out_pmodel$ci + 2 * out_pmodel$gammastar ))
```
Yes.

### Elevation and pressure

Above, atmospheric pressure (`patm`) was not provided as an argument, but elevation (`elv`) was. Hence the warning was printed (only when `verbose = TRUE`), saying: `Atmospheric pressure (patm) not provided. Calculating it as a function of elevation (elv),` `Assuming standard atmosphere (101325 Pa at sea level).`. Alternatively, we can provide atmospheric pressure (`patm`) as input, which overrides the argument `elv`.

### P-model for time series

The `rpmodel()` function can also be invoked for time series, where `tc`, `vpd`, `co2`, `fapar`, `patm`, and `ppfd` are vectors.
```{r}
out_ts_pmodel <- rpmodel( 
  tc             = 20 + rnorm(5, mean = 0, sd = 5),
  vpd            = 1000 + rnorm(5, mean = 0, sd = 50),
  co2            = rep(400, 5),
  fapar          = rep(1, 5),
  ppfd           = 300 + rnorm(5, mean = 0, sd = 30),
  elv            = 0,         
  kphio          = 0.05,         
  beta           = 146,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE,
  verbose        = FALSE
  )
print(out_ts_pmodel$gpp)
```
Note that `gpp` (as well as all other returned variables) are now vectors of the same length as the vectors provided as inputs.

### P-model in the tidyverse

We can create a data frame (in [tidyverse](https://www.tidyverse.org/) this is a tibble) and apply the `rpmodel()` function to each row.
```{r}
library(dplyr)
library(purrr)
df <- tibble(
  tc             = 20 + rnorm(5, mean = 0, sd = 5),
  vpd            = 1000 + rnorm(5, mean = 0, sd = 50),
  co2            = rep(400, 5),
  fapar          = rep(1, 5),
  ppfd           = 300 + rnorm(5, mean = 0, sd = 30)
  ) %>%
  mutate( out_pmodel = purrr::pmap(., rpmodel, 
    elv            = 0,         
    kphio          = 0.05,         
    beta           = 146,
    method_optci   = "prentice14",
    method_jmaxlim = "none",
    do_ftemp_kphio = FALSE
    ) )
print(df)
```
Note that the new column `out_pmodel` now contains the list returned as output of the `rpmodel()` function applied to each row separately. Additional (constant) arguments are just passed to `purrr::pmap` as arguments.

If you prefer the elements of these lists to be in separate columns of `df`, use tidyr to do:
```{r}
library(tidyr)
df <- df %>% 
  mutate( out_pmodel = purrr::map(out_pmodel, ~as_tibble(.))) %>% 
  unnest(out_pmodel)
print(df)
```

## Auxiliary functions

A number of auxiliary functions, which are used within `rpmodel()`, are available (public) through the package. 

### Instantaneous temperature scaling

Different instantaneous temperature scaling functions are applied for $V_\text{cmax}$ and dark respiration ($R_d$). 

- `calc_ftemp_inst_vcmax()` calculates the instantaneous temperature response of $V_\text{cmax}$. Let's run the P-model for `tc = 10` (degrees C). The ratio of $V_\text{cmax}/V_\text{cmax25}$ should equal the instantaneous temperature scaling function for $V_\text{cmax}$ at 10 degrees C (calculated by `calc_ftemp_inst_vcmax(10)`):
```{r}
out_pmodel <- rpmodel( 
  tc             = 10,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 300,          # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  kphio          = 0.05,         # quantum yield efficiency,
  beta           = 146,          # unit cost ratio a/b,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE,
  verbose        = TRUE
  )
print(paste("Ratio Vcmax/Vcmax25      :", out_pmodel$vcmax/out_pmodel$vcmax25))
print(paste("calc_ftemp_inst_vcmax(10):", calc_ftemp_inst_vcmax(10)))
```

- `calc_ftemp_arrh()` Calculates the Arrhenius-type temperature response and is used inside `calc_ftemp_inst_vcmax()`.

- `calc_ftemp_inst_rd()` calculates the temperature response of dark respiration ($R_d$), which is slightly less steep than that for $V_\text{cmax}$:
```{r}
print(paste("calc_ftemp_inst_rd(10):", calc_ftemp_inst_rd(10)))
```

### Parameters in the FvCB model

- `calc_gammastar()` calculates the CO$_2$ compensation point ($\Gamma^\ast$) in the Farquhar-von Caemmerer-Berry model as a function of temperature (argument `tc`) and atmospheric pressure (argument `patm`). This is returned by the `rpmodel()` function and by the separate auxiliary function `calc_gammastar()`. `calc_gammastar()` requires atmospheric pressure (`patm`) to be given as an argument (in addition to temperature). Corresponding to the `rpmodel()` call above, let's calculate this using the auxiliary function `calc_patm()` with 0 metres above sea level, and assuming standard atmospheric pressure (101325 Pa at 0 m a.s.l.):
```{r}
print(paste("From rpmodel call :", out_pmodel$gammastar))
print(paste("calc_gammastar(10):", calc_gammastar(10, patm = calc_patm(elv = 0))))
```

- `calc_kmm()` calculates the Michaelis Menten coefficient for Rubisco-limited photosynthesis as a function of temperature (argument `tc`) and atmospheric pressure (argument `patm`). As above,  `calc_kmm()` requires atmospheric pressure to be given as an argument (in addition to temperature). Corresponding to the `rpmodel()` call above, let's calculate this using the auxiliary function `calc_patm()` with 0 metres above sea level, and assuming standard atmospheric pressure (101325 Pa at 0 m a.s.l.):
```{r}
print(paste("From rpmodel call:", out_pmodel$kmm))
print(paste("calc_kmm(10)     :", calc_kmm(10, patm = calc_patm(elv = 0))))
```

### Temperature dependence of quantum yield efficiency

The temperature dependence of quantum yield efficiency is modelled following Bernacchi et al. (2003), if the argument to the `rpmodel()` call `do_ftemp_kphio = TRUE`. This affects several quantities returned by the `rpmodel()` call (GPP, LUE, Vcmax), and can be calculated direction using `calc_ftemp_kphio()`.
```{r}
out_pmodel_ftemp_kphio_ON <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 300,          # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  do_ftemp_kphio = TRUE
  )
out_pmodel_ftemp_kphio_OFF <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 300,          # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  do_ftemp_kphio = FALSE
  )
print(paste("LUE ftemp_ON /LUE ftemp_OFF =", out_pmodel_ftemp_kphio_ON$lue / out_pmodel_ftemp_kphio_OFF$lue))
print(paste("GPP ftemp_ON /GPP ftemp_OFF =", out_pmodel_ftemp_kphio_ON$gpp / out_pmodel_ftemp_kphio_OFF$gpp))
print(paste("Vcmax ftemp_ON /Vcmax ftemp_OFF =", out_pmodel_ftemp_kphio_ON$vcmax / out_pmodel_ftemp_kphio_OFF$vcmax))
print(paste("calc_ftemp_kphio(20) =", calc_ftemp_kphio(20)))
```


### Soil moisture stress

Similar to above (), the soil moisture dependence of LUE (and hence GPP, and Vcmax) can be calculated directly using the function `calc_soilmstress()` and affects several quantities returned by the `rpmodel()` call (GPP, LUE, Vcmax):
```{r}
out_pmodel_soilmstress_OFF <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 300,          # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  do_ftemp_kphio = FALSE,
  do_soilmstress = TRUE
  )
out_pmodel_soilmstress_ON <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 300,          # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  do_ftemp_kphio = FALSE,
  do_soilmstress = TRUE,
  soilm          = 0.2,
  apar_soilm     = 0.1,
  bpar_soilm     = 0.7,
  meanalpha      = 0.2 
  )
print(paste("LUE soilmstress_ON /LUE soilmstress_OFF =", out_pmodel_soilmstress_ON$lue / out_pmodel_soilmstress_OFF$lue))
print(paste("GPP soilmstress_ON /GPP soilmstress_OFF =", out_pmodel_soilmstress_ON$gpp / out_pmodel_soilmstress_OFF$gpp))
print(paste("Vcmax soilmstress_ON /Vcmax soilmstress_OFF =", out_pmodel_soilmstress_ON$vcmax / out_pmodel_soilmstress_OFF$vcmax))
print(paste("calc_soilmstress(0.2, apar_soilm = 0.1, bpar_soilm = 0.7, meanalpha = 0.2) =", calc_soilmstress(0.2, apar_soilm = 0.1, bpar_soilm = 0.7, meanalpha = 0.2)))
```

`calc_ftemp_arrh()` Calculates the Arrhenius-type temperature response. 

<!-- ## Response to CO2 -->

<!-- ```{r} -->
<!-- library(ggplot) -->
<!-- df <- tibble( -->
<!--   tc             = rep(20, 100), -->
<!--   vpd            = rep(1000, 100), -->
<!--   co2            = seq(360, 1000, length.out = 100), -->
<!--   fapar          = rep(1, 100), -->
<!--   ppfd           = rep(100, 100) -->
<!--   ) %>% -->
<!--   dplyr::mutate( out_pmodel_simpl = purrr::pmap(., rpmodel, -->
<!--     elv            = 0, -->
<!--     kphio          = 0.05, -->
<!--     beta           = 146, -->
<!--     method_optci   = "prentice14", -->
<!--     method_jmaxlim = "none", -->
<!--     do_ftemp_kphio = FALSE -->
<!--     ) ) %>% -->
<!--   dplyr::mutate( out_pmodel_full = purrr::pmap(dplyr::select(., 1:5), rpmodel,  -->
<!--     elv            = 0,          -->
<!--     kphio          = 0.05,          -->
<!--     beta           = 146, -->
<!--     method_optci   = "prentice14", -->
<!--     method_jmaxlim = "wang17", -->
<!--     do_ftemp_kphio = FALSE -->
<!--     ) ) %>%  -->
<!--   dplyr::mutate( out_pmodel_simpl = purrr::map(out_pmodel_simpl, ~as_tibble(.))) %>% -->
<!--   dplyr::mutate( out_pmodel_full  = purrr::map(out_pmodel_full,  ~as_tibble(.))) -->

<!-- df %>%  -->
<!--   tidyr::unnest(out_pmodel_full) %>%  -->
<!--   dplyr::select(co2, out_pmodel_simpl, gpp_full = gpp) %>%  -->
<!--   tidyr::unnest(out_pmodel_simpl) %>%  -->
<!--   dplyr::select(co2, gpp_full, gpp_simpl = gpp) %>%  -->
<!--   tidyr::gather(method, gpp, c(gpp_full, gpp_simpl)) %>%  -->
<!--   ggplot(aes(x = co2, y = gpp, color = method)) + -->
<!--   geom_line() -->

<!-- ``` -->

## References

Stocker, B. D., Wang, H., Smith, N. G., Harrison, S. P., Keenan, T. F., Sandoval, D., Davis, T., and Prentice, I. C.: P-model v1.0: An optimality-based light use efficiency model for simulating ecosystem gross primary production, Geosci. Model Dev. Discuss., https://doi.org/10.5194/gmd-2019-200, in review, 2019.


