---
title: "Theory"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
---

The P-model predicts an optimal ratio of $c_i : c_a$, termed as $\chi$, that balances the costs associated with maintaining the transpiration ($E$) stream and the carboxylation capacity $V_{\text{cmax}}$. It can therefore be used to simulate the acclimation of the photosynthetic machinery to its environment - a mechanism that happens at a time scale of several days to months. At its core, it provides a solution for the optimality criterium
$$
a \; \frac{\partial (E/A)}{\partial \chi} = -b \; \frac{\partial (V_{\mathrm{cmax}}/A)}{\partial \chi}  \;\;\;\;\;\;\;\;\;\;\;\;(1)
$$
The optimal $\chi$ solves the above equation and, with $E = 1.6 g_s D$, $A = g_s (1-\chi)$, and using the Rubisco-limited assimilation rate:
$$
A = A_C = V_{\mathrm{cmax}} \; \frac{\chi\;c_a-\Gamma^{\ast}}{\chi\;c_a + K}
$$
is given by:
$$
\chi = \frac{\Gamma^{\ast}}{c_a} + \left(1- \frac{\Gamma^{\ast}}{c_a}\right)\;\frac{\xi}{\xi + \sqrt{D}}
$$
with
$$
\xi = \sqrt{\frac{b(K+\Gamma^{\ast})}{1.6\;a}}
$$
The unit cost ratio $b/a$ is also referred to as $\beta$.

## P-model run

So much for the theory. Let's run the P-model, without $J_{\text{max}}$ limitation, for one set of inputs, being temperature, PPFD, VPD, CO$_2$, elevation, and fAPAR.

To do so, run the `rpmodel()` function from the rsofun package:
```{r, message=FALSE, warning=FALSE}
library(rpmodel)
library(dplyr)
# modified seq() function to get a logarithmically spaced sequence
lseq <- function(from=1, to=100000, length.out=6) {
  exp(seq(log(from), log(to), length.out = length.out))
}

## Set parameters
beta <- 146          # unit cost ratio a/b
gamma <- 0.105       # unit cost ratio c/b
kphio <- 0.05        # quantum yield efficiency
c_molmass <- 12.0107 # molar mass, g / mol

## Define environmental conditions
tc <- 20             # temperature, deg C
ppfd <- 300          # mol/m2/d
vpd  <- 1000          # Pa
co2  <- 400          # ppm
elv  <- 0            # m.a.s.l.
fapar <- 1           # fraction

out_analytical <- rsofun::rpmodel(
  tc             = tc,
  vpd            = vpd,
  co2            = co2,
  elv            = elv,
  kphio          = kphio,
  beta           = beta,
  fapar          = fapar,
  ppfd           = ppfd,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE
  )
```

The function returns a list of variables (see also man page by `?rpmodel`), including $V_{\mathrm{cmax}}$, $g_s$, and all the parameters of the photosynthesis model ($K$, $\Gamma^{\ast}$), which are all internally consistent, as can be verified for...
$$
c_i = c_a - A / g_s = \chi c_a
$$

```{r}
print( out_analytical$ci )
print( out_analytical$ca - (out_analytical$gpp / c_molmass) / out_analytical$gs )
print( out_analytical$ca * out_analytical$chi )
```
Yes.

And for...
$$
A = V_{\text{cmax}} \frac{c_i-\Gamma^{\ast}}{c_i + K} = \phi_0 I_{\text{abs}} \frac{c_i-\Gamma^{\ast}}{c_i + 2 \Gamma^{\ast}} = g_s (c_a - c_i)
$$

```{r}
print( out_analytical$gpp / c_molmass )
print( out_analytical$vcmax * (out_analytical$ci - out_analytical$gammastar) / (out_analytical$ci + out_analytical$kmm ))
print( out_analytical$gs * (out_analytical$ca - out_analytical$ci) )

print( kphio * ppfd * fapar * (out_analytical$ci - out_analytical$gammastar) / (out_analytical$ci + 2 * out_analytical$gammastar ))
```
Yes.

## Numerical solution

Instead of formulating the optimality criterium (Eq. 1) with respect to equality in marginal costs (derivative w.r.t. $\chi$), one may also write this as a minimisation of the combined costs. With $\beta=b/a$, the following criterium is equivalent to Eq. 1:
$$
E/A + \beta V_{\mathrm{cmax}}/A = min. \;\;\;\;\;\;\;\;(2)
$$
This formulation has the advantage that we can easily apply a numerial search algorithm to find the solution in $\chi$. Why bother? I think it has an advantage as it allows for a more modular treatment, e.g., of transpiration as a function of $g_s$. But for now, let's stick with $E=1.6g_sD$. Eq. 2 can be written out as:
$$
\frac{1.6\;D}{c_a(1-\chi)} + \beta \frac{\chi c_a + K}{\chi c_a - \Gamma^{\ast}} = min.
$$

To find the minimum, let's define the numerical optimization function using the L-BFGS-B algorithm implemented in the `optimr` package:
```{r}
calc_optimal_chi_num <- function( kmm, gammastar, ns_star, ca, vpd, beta ){
  #-----------------------------------------------------------------------
  # Input:    - float, 'kmm' : Pa, Michaelis-Menten coeff.
  #           - float, 'ns_star'  : (unitless) viscosity correction factor for water
  #           - float, 'vpd' : Pa, vapor pressure deficit
  # Output:   float, ratio of ci/ca (chi)
  # Features: Returns an estimate of leaf internal to ambient CO2
  #           partial pressure following the "simple formulation".
  # Depends:  - kc
  #           - ns
  #           - vpd
  #-----------------------------------------------------------------------
  maximise_this <- function( chi, kmm, gammastar, ns_star, ca, vpd, beta ){
    out <- 1.6 * ns_star * vpd / (ca * (1.0 - chi)) + beta * (chi * ca + kmm)/(chi * ca - gammastar)
    return(out)
  }

  out_optim <- optimr::optimr(
    par       = 0.7,
    lower     = 0.1,
    upper     = 0.99,
    fn        = maximise_this,
    kmm       = kmm,
    gammastar = gammastar,
    ns_star   = ns_star,
    ca        = ca,
    vpd       = vpd,
    beta      = beta,
    method    = "L-BFGS-B",
    control   = list( maxit = 100, maximize = TRUE )
    )

  return(out_optim$par)
}
```

Now, let's find the optimal $\chi$ numerically using above function with the same parameters as calculated inside and returned by the P-model.
```{r}
chi_opt <- calc_optimal_chi_num( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta )
print(chi_opt)
```
As a check: is this the same as returned by the P-model?
```{r}
print(out_analytical$chi)
```
Practically, yes. A slight difference is due to the limited number of iterations (`maxit=100`).

Let's call this $\chi^{\ast}$. For the P-model, it is then assumed that $A=A_J$ with
$$
A_J = \phi_0 \; I_{\mathrm{abs}}\;\frac{c_a \chi^{\ast} - \Gamma^{\ast}}{c_a \chi^{\ast} + 2\Gamma^{\ast}}
$$
It is further assumed that the light-limited and Rubisco-limited assimilation rates are equal for average conditions (which implies that the arguments to the P-model run should represent average conditions). This assumption is also called the "coordination hypothesis."
$$
A_J = A_C
$$
This allows for solving for $V_{\text{cmax}}$:
$$
V_{\text{cmax}} = \phi_0 \; I_{\mathrm{abs}}\;\frac{c_a \chi^{\ast} + K}{c_a \chi^{\ast} + 2 \Gamma^{\ast} }
$$
This is what the P-model returns as can be verified:
```{r}
print( out_analytical$vcmax)
print( kphio * fapar * ppfd * (out_analytical$ci + out_analytical$kmm) / (out_analytical$ci + 2 * out_analytical$gammastar ) )
```

### Performance benchmark

TODO: Check computation time for N calls of the numerical `calc_optimal_chi_num()` vs. the analytical `calc_optimal_chi()` functions.


## Optimising $g_s$ and $V_{\text{cmax}}$ instead of $\chi$

At the core of the assimilation-transpiration trade-off are the quantities $g_s$ and $V_{\text{cmax}}$. The combined effect of their magnitudes determines $\chi$. Why not simutaneously optimize $g_s$ and $V_{\text{cmax}}$? We can write Eq. 2 also as:
$$
(1.6g_sD + \beta V_{\text{cmax}})/A_C = min. \;\;\;\;\;\;\;(3)
$$
and implement this in another numerical search function. Note that $c_i$ can be calculated from solving the equation system
$$
A = V_{\mathrm{cmax}} \; \frac{c_i-\Gamma^{\ast}}{c_i + K} \\
A = g_s(c_a - c_i)
$$
which leads to a quadratic equation for $c_i$ where we will only take the positive real part (`Re(root_ci)[which(Re(root_ci)>0)]`). This can be implemented as follows.
```{r}
calc_optimal_gs_vcmax <- function( par, args, maximize=FALSE, return_all=FALSE ){

  kmm       <- args[1]
  gammastar <- args[2]
  ns_star   <- args[3]
  ca        <- args[4]
  vpd       <- args[5]
  beta      <- args[6]

  vcmax <- par[1]
  gs    <- par[2]

  ## Get assimilation and ci, given Vcmax and gs
  ## by solving the equation system
  ## assim = vcmax * (ci - gammastar)/(ci + kmm)
  ## assim = gs * (ca - ci)
  a_quad <- -1.0 * gs
  b_quad <- gs * ca - gs * kmm - vcmax
  c_quad <- gs * ca * kmm + vcmax * gammastar

  root_ci <- try( polyroot( c(c_quad, b_quad, a_quad) ) )

  if (class(root_ci)=="try-error"){

    return( list( vcmax=NA, gs=NA, ci=NA, chi=NA, a_c=NA, cost_transp=NA, cost_vcmax=NA, net_assim=NA  ) )

  } else {

    ## take only real part of the root
    ci <- Re(root_ci)

    ## take only positive root
    ci <- ci[which(ci>0)]   # take positive root

    ## if both are positive, take the one that satisfies ci < ca (don't know if this is necessary)
    if (length(ci)>1) ci <- ci[which(ci<ca)]

    ## A_c
    a_c <- vcmax * (ci - gammastar) / (ci + kmm)

    ## only cost ratio is defined. for this here we need absolute values. Set randomly
    cost_transp <- 1.6 * ns_star * gs * vpd
    cost_vcmax  <- beta * vcmax

    ## Option B: This is equivalent to the P-model with its optimization of ci:ca.
    net_assim <- -(cost_transp + cost_vcmax) / a_c

    if (maximize) net_assim <- -net_assim

    if (return_all){
      return( list( vcmax=vcmax, gs=gs, ci=ci, chi=ci/ca, a_c=a_c, cost_transp=cost_transp, cost_vcmax=cost_vcmax, net_assim=net_assim  ) )
    } else {
      return( net_assim )
    }

  }
}
```

Now that the target function is defined, we can actually numerically search for the optimal $V_{\text{cmax}}$ and $g_s$. We need to provide the algorithm with starting values, taken here as the $V_{\text{cmax}}$ and $g_s$ returned by our initial P-model call.
```{r}
wrap_calc_optimal_gs_vcmax <- function( kmm, gammastar, ns_star, ca, vpd, beta, vcmax_start, gs_start ){
  out_optim <- optimr::optimr(
    par        = c( out_analytical$vcmax, out_analytical$gs ), # starting values
    lower      = c( out_analytical$vcmax*0.0001, out_analytical$gs*0.001 ),
    upper      = c( out_analytical$vcmax*20,   out_analytical$gs*30 ),
    fn         = calc_optimal_gs_vcmax,
    args       = c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta ),
    method     = "L-BFGS-B",
    maximize   = TRUE,
    control    = list( maxit = 1000 )
    )
  varlist <- calc_optimal_gs_vcmax( par=out_optim$par, args=c(kmm, gammastar, ns_star, ca, vpd, beta), maximize=FALSE, return_all=TRUE )
  return(varlist)
}

```

Note that, given that the optimality criterium implemented here (Eq. 3) embodies the same idea and is mathematically identical as Eq. 1, this should yield the same $\chi$ (which can be back-calculated from the optimised $V_{\text{cmax}}$ and $g_s$). Let's check if this is the case.
```{r echo=FALSE}
## Invoke function with optimised Vcmax and gs (in out_optim$par), now returning all variables
varlist <- wrap_calc_optimal_gs_vcmax( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta, out_analytical$vcmax, out_analytical$gs )
print( paste("Optimal chi from Vcmax-gs optimization:", varlist$chi ) )
print( paste("Optimal chi from P-model:", out_analytical$chi ) )
```
Indeed, this is the same.

Now, we have an optimised $V_{\text{cmax}}$ (lets call it $V_{\text{cmax}}^{\ast}$), and we can have an optimised $c_i$ which is consistent with the former. Given this, we can take the same step as is done in the P-model and assume that $A_J=A_C$, which leads to
$$
V_{\text{cmax}} = \phi_0 \; I_{\mathrm{abs}}\;\frac{c_a \chi^{\ast} + K}{c_a \chi^{\ast} + 2 \Gamma^{\ast} }
$$
To verify one again:
```{r}
print( paste( "optimal Vcmax:", varlist$vcmax ) )
print( paste( "Vcmax^P:", kphio * fapar * ppfd * (varlist$ci + varlist$kmm) / (varlist$ci + 2 * varlist$gammastar )) )
print( paste("Vcmax from P-model:", out_analytical$vcmax ) )
```

It's important to note that calculating $V_{\text{cmax}}$ like this implies the assumption of $A_J = A_C$, while solving for simultaneously optimal $V_{\text{cmax}}$ and $g_s$ doesn't use this assumption. Nevertheless, the results are identical, which may not be immediately obvious.

### Light limitation?

But what about the implicit assumption made in the optmisation that assimilation always follows the Rubisco-limited curve? Shouldn't there be a limit given by light availability? We can spell this out by calculating $A=min(A_C, A_J)$ or equivalently $c_i=max(c_i^C, c_i^J)$ and modify `calc_optimal_gs_vcmax()` accordingly. We now need to additionally solve the equation system
$$
A = \phi_0 I_{\text{abs}} \; \frac{c_i-\Gamma^{\ast}}{c_i + 2 \Gamma^{\ast}} \\
A = g_s(c_a - c_i)
$$

This is implemented as follows:
```{r}
calc_optimal_gs_vcmax_ll <- function( par, args, iabs, kphio, maximize=FALSE, return_all=FALSE ){

  kmm       <- args[1]
  gammastar <- args[2]
  ns_star   <- args[3]
  ca        <- args[4]
  vpd       <- args[5]
  beta      <- args[6]

  vcmax <- par[1]
  gs    <- par[2]

  ## Get ci using Rubisco-limited assimilation
  ## Get assimilation and ci, given Vcmax and gs
  ## by solving the equation system
  ## assim = vcmax * (ci - gammastar)/(ci + kmm)
  ## assim = gs * (ca - ci)
  a_quad <- -1.0 * gs  # coefficient of quadratic term
  b_quad <- gs * ca - gs * kmm - vcmax
  c_quad <- gs * ca * kmm + vcmax * gammastar

  root_ci_c <- try( polyroot( c(c_quad, b_quad, a_quad) ) )

  ## Get ci using light-limited assimilation
  ## Get assimilation and ci, given Vcmax and gs
  ## by solving the equation system
  ## assim = kphio * iabs * (ci-gammastar)/(ci+2*gammastar)
  ## assim = gs * (ca - ci)
  a_quad <- gs  # coefficient of quadratic term
  b_quad <- kphio * iabs - gs * ca + 2 * gs * gammastar
  c_quad <- - kphio * iabs * gammastar - 2 * gs * ca * gammastar

  root_ci_j <- try( polyroot( c(c_quad, b_quad, a_quad) ) )

  if (class(root_ci_j)=="try-error" || class(root_ci_c)=="try-error"){

    return( list( vcmax=NA, gs=NA, ci=NA, chi=NA, a_c=NA, a_j=NA, ci_c=NA, ci_j=NA, cost_transp=NA, cost_vcmax=NA, net_assim=NA  ) )

  } else {

    ## take only real part of the root
    ci_j <- Re(root_ci_j)
    ci_c <- Re(root_ci_c)

    ## take only positive root
    ci_j <- ci_j[which(ci_j>0)]   # take positive root
    ci_c <- ci_c[which(ci_c>0)]   # take positive root

    ## if both are positive, take the one that satisfies ci < ca (don't know if this is necessary)
    if (length(ci_j)>1) ci_j <- ci_j[which(ci_j<ca)]
    if (length(ci_c)>1) ci_c <- ci_c[which(ci_c<ca)]

    ## Rubisco-limited
    a_c <- vcmax * (ci_c - gammastar) / (ci_c + kmm)

    ## light-limited
    a_j <- kphio * iabs * (ci_j - gammastar) / (ci_j + 2 * gammastar)

    ## Take minimum of the two assimilation rates
    assim <- min( a_j, a_c )

    # if (return_all){
    #   if (a_j<a_c) print("Warning: A_J is lower than A_C")
    # }

    ## ... and consistently the maximum of the two ci
    ci <- max( ci_j, ci_c  )

    ## only cost ratio is defined. for this here we need absolute values. Set randomly
    cost_transp <- 1.6 * ns_star * gs * vpd
    cost_vcmax  <- beta * vcmax

    ## Option B: This is equivalent to the P-model with its optimization of ci:ca.
    net_assim <- -(cost_transp + cost_vcmax) / assim

    if (maximize) net_assim <- -net_assim

    if (return_all){
      return( list( vcmax=vcmax, gs=gs, ci=ci, chi=ci/ca, a_c=a_c, a_j=a_j, ci_c=ci_c, ci_j=ci_j, cost_transp=cost_transp, cost_vcmax=cost_vcmax, net_assim=net_assim  ) )
    } else {
      return( net_assim )
    }

  }
}
```

Let's compare the results for $\chi$ with this modified version to the original one:
```{r}
wrap_calc_optimal_gs_vcmax_ll <- function( ppfd, fapar, kphio, kmm, gammastar, ns_star, ca, vpd, beta, vcmax_start, gs_start ){
  out_optim <- optimr::optimr(
    par        = c( vcmax_start, gs_start ), # starting values
    lower      = c( vcmax_start*0.0001, gs_start*0.001 ),
    upper      = c( vcmax_start*20, gs_start*30 ),
    fn         = calc_optimal_gs_vcmax_ll,
    args       = c(kmm, gammastar, ns_star, ca, vpd, beta),
    iabs       = (ppfd * fapar),
    kphio      = kphio,
    method     = "L-BFGS-B",
    maximize   = TRUE,
    control    = list( maxit = 1000 )
    )
  varlist <- calc_optimal_gs_vcmax_ll( par=out_optim$par, args=c(kmm, gammastar, ns_star, ca, vpd, beta), iabs=(fapar*ppfd), kphio=kphio, maximize=FALSE, return_all=TRUE )
  return(varlist)
}

varlist_ll <- wrap_calc_optimal_gs_vcmax_ll( ppfd, fapar, kphio, out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta, out_analytical$vcmax, out_analytical$gs )

## Invoke function with optimised Vcmax and gs (in out_optim$par), now returning all variables
print( paste("Optimal chi from P-model:", out_analytical$chi ) )
print( paste("Optimal chi from Vcmax-gs optimization:", varlist$chi ) )
print( paste("Optimal chi from Vcmax-gs opt. with light limit.:", varlist_ll$chi ) )
```

In this case, it's not a problem. It yields the same $\chi$. But is this the case across a range of light conditions (remember, that $\chi$ is independent of light in the original model)?
```{r}
ppfd_vec <- seq(from=0, to=8000, length.out=50)
out_opt_ll_list <- purrr::map(
  as.list( ppfd_vec),
  ~wrap_calc_optimal_gs_vcmax_ll( ., fapar, kphio, out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta, out_analytical$vcmax, out_analytical$gs )
  )
chi_vec <- out_opt_ll_list %>% purrr::map_dbl("chi")
plot( ppfd_vec, chi_vec, type="l", xlab="PPFD", ylab="chi")
```

It is the same. But why?

<!-- Oooops. Our optimal $V_{\text{cmax}}$ is not the same. What does this mean? The P-model assumes that assimilation can be calculated form $c_i$ from the light-limited assimilation rate. The trouble is that if the light-limited one is lower than the Rubisco-limited one, then the optimisation derives a $V_{\text{cmax}}$ at which it cannot operate. The $V_{\text{cmax}}$ derived from optimised $\chi$ (and hence $c_i$) is not what was "paid for" in the optmisation. -->


## The same with a twist

Arguably the most intuitive way to specify an optimality criterium that (seemingly) embodies the same idea like Eqs. 1 and 3, is to maximise a "net assimilation" rate, defined as assimilation *minus* the costs for maintaining the transpiration stream and carboxylation capacity, instead of the costs per unit assimilation as done above. We can write this as follows:

Maximise the following term numerically, again subject to $V_{\mathrm{cmax}}$ and $g_s$:
$$
A_n = A - aE -bV_{\text{cmax}} = max. \;\;\;\;\;\;\;\;\;\;\;\;(4)
$$
We can divide this by $a$ in order to define it with the cost ratio $\beta$ which we have used above. However, an absolute cost remains in the first term:
$$
A_n = \frac{1}{a} A - E -\beta V_{\text{cmax}} = max. \;\;\;\;\;\;\;\;\;\;\;\;(5)
$$
So be it. But this has implications because we now need to know the abosulte cost terms and not just their ratio. Anyways, we can implement Eq. 5 analogously to above, now with an additional argument `cost_scalar`.
```{r}
calc_net_assim_eq4 <- function( par, args, iabs, kphio, cost_scalar, maximize=FALSE, return_all=FALSE ){

  kmm       <- args[1]
  gammastar <- args[2]
  ns_star   <- args[3]
  ca        <- args[4]
  vpd       <- args[5]
  beta      <- args[6]

  vcmax <- par[1]
  gs    <- par[2]

  ## Get assimilation and ci, given Vcmax and gs
  ## by solving the equation system
  ## assim = vcmax * (ci - gammastar)/(ci + kmm)
  ## assim = gs * (ca - ci)
  a_quad <- -1.0 * gs
  b_quad <- gs * ca - gs * kmm - vcmax
  c_quad <- gs * ca * kmm + vcmax * gammastar

  root_ci <- try( polyroot( c(c_quad, b_quad, a_quad) ) )

  if (class(root_ci)=="try-error"){

    return( list( vcmax=NA, gs=NA, ci=NA, chi=NA, a_c=NA, cost_transp=NA, cost_vcmax=NA, net_assim=NA ) )

  } else {

    ## take only real part of the root
    ci <- Re(root_ci)

    ## take only positive root
    ci <- ci[which(ci>0)]   # take positive root

    ## if both are positive, take the one that satisfies ci < ca (don't know if this is necessary)
    if (length(ci)>1) ci <- ci[which(ci<ca)]

    ## A_c
    a_c <- vcmax * (ci - gammastar) / (ci + kmm)

    ## only cost ratio is defined. for this here we need absolute values. Set randomly
    ## here cost_scalar basically corresponds to a in Eq. 2
    cost_transp <- cost_scalar * 1.6 * ns_star * gs * vpd
    cost_vcmax  <- cost_scalar * beta * vcmax

    ## The trouble is that A_c is used to determine optimality here, but actual assimilation is
    ## later calculated following the light-limited function for A_j.
    net_assim <- a_c - cost_transp - cost_vcmax

    if (maximize) net_assim <- -net_assim

    if (return_all){
      return( list( vcmax=vcmax, gs=gs, ci=ci, chi=ci/ca, a_c=a_c, cost_transp=cost_transp, cost_vcmax=cost_vcmax, net_assim=net_assim  ) )
    } else {
      return( net_assim )
    }
  }
}
```
We don't know `cost_scalar`, so let's just pick a random one (0.0003 used here) and hope we get the same result for $\chi$ as above.
```{r echo=FALSE}
out_optim <- optimr::optimr(
  par        = c( out_analytical$vcmax, out_analytical$gs ), # starting values
  lower      = c( out_analytical$vcmax*0.0001, out_analytical$gs*0.001 ),
  upper      = c( out_analytical$vcmax*20,   out_analytical$gs*30 ),
  fn         = calc_net_assim_eq4,
  args       = c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta ),
  iabs       = (fapar * ppfd),
  kphio      = kphio,
  cost_scalar = 0.0003,
  method     = "L-BFGS-B",
  maximize   = TRUE,
  control    = list( maxit = 1000 )
  )
print( paste( "optimal Vcmax:", out_optim$par[1] ) )
print( paste( "optimal gs:", out_optim$par[2] ) )

## Invoke function with optimised Vcmax and gs (in out_optim$par), now returning all variables
varlist <- calc_net_assim_eq4( par=out_optim$par, args=c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta ), iabs=(fapar * ppfd), kphio=kphio, cost_scalar=0.0003, maximize=FALSE, return_all=TRUE )
print( paste("Optimal chi from Vcmax-gs optimization (Eq. 5):", varlist$chi ) )
print( paste("Optimal chi from P-model:", out_analytical$chi ) )
```
Ooops. Not the same.

The problem of all of this is that the cost factor that scales both the cost of $V_{\text{cmax}}$ and $g_s$ equally is missing from the optimization criterium when $E/A + \beta V_{\text{cmax}}/A = max.$ is used.
Let's try to find a value for the cost scalar so that we get the same $\chi$ as when following Eq. 2 or 3 for the opmitisation.
```{r message=FALSE, warning=FALSE}
setzero <- function( x_cost, chi_target, out_analytical, fapar, ppfd, vpd ){

  args <- c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta )

  out_optim <- optimr::optimr(
    par        = c( out_analytical$vcmax, out_analytical$gs ),
    lower      = c( out_analytical$vcmax*0.0001, out_analytical$gs*0.001 ),
    upper      = c( out_analytical$vcmax*20,   out_analytical$gs*30 ),
    fn         = calc_net_assim_eq4,
    args       = args,
    iabs       = (fapar * ppfd),
    kphio      = kphio,
    cost_scalar = x_cost,
    maximize   = TRUE,
    method     = "L-BFGS-B",
    control    = list( maxit = 1000 )
    )

  list_numerical <- calc_net_assim_eq4( out_optim$par, args, iabs=(fapar * ppfd), kphio = kphio, cost_scalar=x_cost, return_all = TRUE )

  out <- chi_target - list_numerical$chi

  return(abs(out))
}

sapply( lseq( from=0.0001, to=0.005, length.out = 100), FUN =  function(x) setzero( x, out_analytical$chi, out_analytical, fapar, ppfd, vpd ) ) %>%
plot( lseq( from=0.0001, to=0.005, length.out = 100), ., type="l" )
abline(h=0, lty=2)

## Watch out: numerical noise is happening. For this example, chose starting point where visually determined in plot above
unitcost_root <- optimr::optimr(
  par            = 0.002,
  fn             = setzero,
  chi_target     = out_analytical$chi,
  out_analytical = out_analytical,
  fapar          = fapar,
  ppfd           = ppfd,
  vpd            = vpd,
  control        = list( maxit = 100000 )
  )
abline( v = unitcost_root$par, col="red" )
```

Hence, the cost scalar for which the result for $\chi$ is is identical to the P-model's solution, is:
```{r}
print( unitcost_root$par )
```

Let's verify this as above, setting `cost_scalar = unitcost_root$par`:
```{r echo=FALSE}
out_optim <- optimr::optimr(
  par        = c( out_analytical$vcmax, out_analytical$gs ), # starting values
  lower      = c( out_analytical$vcmax*0.0001, out_analytical$gs*0.001 ),
  upper      = c( out_analytical$vcmax*20,   out_analytical$gs*30 ),
  fn         = calc_net_assim_eq4,
  args       = c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta ),
  iabs       = (fapar * ppfd),
  kphio      = kphio,
  cost_scalar = unitcost_root$par,
  method     = "L-BFGS-B",
  maximize   = TRUE,
  control    = list( maxit = 1000 )
  )

## Invoke function with optimised Vcmax and gs (in out_optim$par), now returning all variables
varlist <- calc_net_assim_eq4( par=out_optim$par, args=c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta ), iabs=(fapar * ppfd), kphio=kphio, cost_scalar=unitcost_root$par, maximize=FALSE, return_all=TRUE )
print( paste("Optimal chi from absolute-cost-Vcmax-gs optimization (Eq. 5):", varlist$chi ) )
print( paste("Optimal chi from P-model:", out_analytical$chi ) )
```

Ok, they are now identical.

Some illustration of this:
Plot a surface of $A_n$ as a function of $V_{\text{cmax}}$ and $g_s$, with a red point where the numerical optimization found the maximum, and a green point, where the analytical solution lies.
```{r, message=FALSE, warning=FALSE}
len_vcmax <- 50
len_gs <- 50
vcmax_vec <- lseq(out_analytical$vcmax*0.0001, out_analytical$vcmax*5, length.out = len_vcmax)
gs_vec <- lseq(out_analytical$gs*0.001, out_analytical$gs*5, length.out = len_gs)
args <- c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta )

A_net <- expand.grid(vcmax_vec, gs_vec) %>%
  setNames( c("vcmax", "gs") ) %>%
  rowwise() %>%
  do( net_assim = calc_net_assim_eq4(
    c(.$vcmax, .$gs),
    args=args,
    iabs=(fapar * ppfd),
    kphio = kphio,
    cost_scalar=unitcost_root$par,
    maximize=FALSE
    ) ) %>%
  tidyr::unnest(net_assim) %>%
  as.matrix() %>%
  array( dim=c(len_vcmax,len_gs) )

pmat <- persp( (vcmax_vec), (gs_vec), A_net,
               phi = 10, # Azimut, 0 is along gs_vec axis
               theta = 30,
               col="lightblue"
               )
points( trans3d( (out_optim$par[1]), (out_optim$par[2]), -out_optim$value, pmat ), pch=16, col="red", cex=1.8 )

# Add point for analytically-determined Vcmax and gs, and the net assimilation rate with the same unit cost as above:
cost_transp <- unitcost_root$par * 1.6 * out_analytical$ns_star * out_analytical$gs * vpd
cost_vcmax <- beta * unitcost_root$par * out_analytical$vcmax
net_assim_analytical <- out_analytical$gpp / c_molmass - cost_transp - cost_vcmax
points( trans3d( (out_analytical$vcmax), (out_analytical$gs), net_assim_analytical, pmat ), pch=16, col="green", cex=1 )

# numerical (of course, the value is not the same as analytical, because it's the optimization target and not A_j. Re-calculate A_j from optimized Vcmax and gs.):
print( c(out_optim$par[1], out_optim$par[2], -out_optim$value) )
out_optim_numerical <- calc_net_assim_eq4(
  par            = c(out_optim$par[1], out_optim$par[2]),
  args           = c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta ),
  iabs           = (fapar * ppfd),
  kphio          =  kphio,
  cost_scalar    = unitcost_root$par,
  return_all     = TRUE
  )
print( c( out_optim_numerical$vcmax, out_optim_numerical$gs, out_optim_numerical$net_assim ))
print( c(out_analytical$vcmax, out_analytical$gs, net_assim_analytical) )
```

And in 2D. For fixed (optimal) $g_s$.
```{r}
## cost_scalar is searched by hand
sapply( as.list(vcmax_vec), FUN = function(x) calc_net_assim_eq4( c(x, out_optim$par[2]), args, iabs=(fapar * ppfd), kphio = kphio, cost_scalar = unitcost_root$par ) ) %>%
  plot( vcmax_vec, ., type="l", xlab="Vcmax", ylab="net A" )
abline(v=out_optim$par[1], lty=3)
points( (out_optim$par[1]), -out_optim$value, pch=16, col="red", cex=1.8 )
points( (out_analytical$vcmax), net_assim_analytical, pch=16, col="green", cex=1 )
```

And for fixed (optimal) $V_{\text{cmax}}$:
```{r}
## cost_scalar is searched by hand
sapply( as.list(gs_vec), FUN = function(x) calc_net_assim_eq4( c(out_optim$par[1], x), args, iabs=(fapar * ppfd), kphio = kphio, cost_scalar = unitcost_root$par ) ) %>% plot( gs_vec, ., type="l", xlab="gs", ylab="net A" )
abline( v=out_optim$par[2], lty=3 )
points( (out_optim$par[2]), -out_optim$value, pch=16, col="red", cex=1.8 )
points( (out_analytical$gs), net_assim_analytical, pch=16, col="green", cex=1 )
```

<!-- The reason why the red and green dots are not exactly on top of each other is because the optimisation (red) returns $A_{n}=A_c - \psi_{\text{transp}} - \psi_{\text{Vcmax}}$, while the analytical (green point) returns $A_n = A_j - \psi_{\text{transp}} - \psi_{\text{Vcmax}}$. -->

## Optimality for the diurnal cycle

Variability in ambient conditions is small between days (within a limited period, typically weeks-months), compared to conditions within days. Non-linearity in $A_J$ and $A_C$, and $\chi$ as a function of $D$ suggests that using average daytime conditions is not a valid simplification for optimality in $\chi$ for an entire diurnal cycle (assuming that $\chi$, and hence $V_{\text{cmax}}$ and $g_s$, don't vary throughout a day). 

What we're actually after is:
$$
\sum_t  (A_t - a E_t - bV_{\text{cmax}, t}) = max. \;\;\;\;\;\;\;\;(6)
$$
This can also be expressed as:
$$
a \frac{\sum E_t}{\sum A_t} + b \frac{\sum V_{\text{cmax}, t}}{\sum A_t} = min. \;\;\;\;\;\;\;\;(7)
$$
A minimum is given when the derivative of this expression w.r.t. $\chi$ is zero:
$$
a \frac{\partial \frac{\sum E}{\sum A}}{\partial \chi} + b \frac{\partial \frac{\sum V_{\text{cmax}}}{\sum A}}{\partial \chi} = 0  \;\;\;\;\;\;\;\;(8)
$$
The trouble is that 
$$
\frac{\sum E}{\sum A} \ne \sum E/A
$$
and hence this Eq. 8 cannot be solved analytically (or can it?). Instead, one could solve Eq. 7 numerically.


## Introducing $J_{\text{max}}$ limitation

From Eq. 4 it's straight-forward to introduce the cost of maintaining a certain $J_{\text{max}}$ in the optimality criterium as
$$
A_n = A - aE -bV_{\text{cmax}} - cJ_{\text{max}} = max. \;\;\;\;\;\;\;\;\;\;\;\;(5)
$$

In the optimisation routine it is now essential to take $A = min(A_C, A_J)$. With $J_{\text{max}}$ limitation factored in using the Smith (1937) formulation, $A_J$ is
$$
A_J = \phi_0 \; I_{\mathrm{abs}} \; \frac{c_i - \Gamma^{\ast}}{c_i + 2\Gamma^{\ast}} \; \frac{1}{\sqrt{1+ \left( \frac{4\;\phi_0\;I_{\mathrm{abs}}}{J_{\mathrm{max}}} \right)^{2}}}
$$

We can implement this as follows::
```{r}
calc_optimal_gs_vcmax_jmax <- function( par, args, iabs, kphio, cost_scalar, maximize=FALSE, return_all=FALSE ){

  kmm       <- args[1]
  gammastar <- args[2]
  ns_star   <- args[3]
  ca        <- args[4]
  vpd       <- args[5]
  beta      <- args[6]
  gamma <- args[7]

  vcmax <- par[1]
  gs    <- par[2]
  jmax  <- par[3]

  calc_L_factor <- function( jmax, iabs, kphio ){
    1.0 / sqrt( 1 + (4 * kphio * iabs / jmax )^2 )
  }

  ## Get ci using Rubisco-limited assimilation
  ## Get assimilation and ci, given Vcmax and gs
  ## by solving the equation system
  ## assim = vcmax * (ci - gammastar)/(ci + kmm)
  ## assim = gs * (ca - ci)
  a_quad <- -1.0 * gs  # coefficient of quadratic term
  b_quad <- gs * ca - gs * kmm - vcmax
  c_quad <- gs * ca * kmm + vcmax * gammastar

  root_ci_c <- try( polyroot( c(c_quad, b_quad, a_quad) ) )

  ## Get ci using light-limited assimilation
  ## Get assimilation and ci, given Vcmax and gs
  ## by solving the equation system
  ## assim = kphio * iabs * (ci-gammastar)/(ci+2*gammastar)
  ## assim = gs * (ca - ci)
  L_factor <- calc_L_factor( jmax, iabs, kphio )
  a_quad <- L_factor * gs  # coefficient of quadratic term
  b_quad <- kphio * iabs - L_factor * gs * ca + 2 * L_factor * gs * gammastar
  c_quad <- - kphio * iabs * gammastar - 2 * L_factor * gs * ca * gammastar    # minus was missing!!!

  root_ci_j <- try( polyroot( c(c_quad, b_quad, a_quad) ) )

  if (class(root_ci_j)=="try-error" || class(root_ci_c)=="try-error"){

    if (!return_all){
      return(9999.9)
    } else {
      return( list( vcmax=NA, gs=NA, ci=NA, chi=NA, a_c=NA, a_j=NA, ci_c=NA, ci_j=NA, cost_transp=NA, cost_vcmax=NA, cost_jmax=NA, net_assim=NA ) )
    }

  } else {

    ## take only real part of the root
    ci_j <- Re(root_ci_j)
    ci_c <- Re(root_ci_c)

    ## take only positive root
    ci_j <- ci_j[which(ci_j>0)]   # take positive root
    ci_c <- ci_c[which(ci_c>0)]   # take positive root

    ## if both are positive, take the one that satisfies ci < ca (don't know if this is necessary)
    if (length(ci_j)>1) ci_j <- ci_j[which(ci_j<ca)]
    if (length(ci_c)>1) ci_c <- ci_c[which(ci_c<ca)]

    ## Rubisco-limited
    a_c <- vcmax * (ci_c - gammastar) / (ci_c + kmm)

    ## light-limited
    a_j <- kphio * iabs * (ci_j - gammastar) / (ci_j + 2 * gammastar)

    ## Take minimum of the two assimilation rates
    assim <- min( a_j, a_c )

    # if (return_all){
    #   if (a_j<a_c) print("Warning: A_J is lower than A_C")
    # }

    ## ... and consistently the maximum of the two ci
    ci <- max( ci_j, ci_c  )

    ## only cost ratio is defined. for this here we need absolute values. Set randomly
    ## here cost_scalar basically corresponds to a in Eq. 2
    cost_transp <- cost_scalar * 1.6 * ns_star * gs * vpd
    cost_vcmax  <- cost_scalar * beta * vcmax
    cost_jmax   <- gamma * jmax    # Tried here without cost_scalar because we know that dA/dJmax = 0.105 = gamma

    ## The trouble is that A_c is used to determine optimality here, but actual assimilation is
    ## later calculated following the light-limited function for A_j.
    net_assim <- assim - cost_transp - cost_vcmax - cost_jmax

    if (maximize) net_assim <- -net_assim

    if (return_all){
      return( list( vcmax=vcmax, gs=gs, ci=ci, chi=ci/ca, a_c=a_c, a_j=a_j, ci_c=ci_c, ci_j=ci_j, cost_transp=cost_transp, cost_vcmax=cost_vcmax, cost_jmax=cost_jmax, net_assim=net_assim  ) )
    } else {
      return( net_assim )
    }

  }
}
```

Let's compare the results for $\chi$ with $J_{\text{max}}$ limitation and without.
```{r}
## test
test <- calc_optimal_gs_vcmax_jmax(
  par = c( out_analytical$vcmax, out_analytical$gs, out_analytical$vcmax ),
  args = c( out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta, gamma),
  iabs = (ppfd * fapar),
  kphio = kphio,
  cost_scalar = unitcost_root$par,
  maximize=FALSE,
  return_all=TRUE
  )

wrap_calc_optimal_gs_vcmax_jmax <- function( ppfd, fapar, kphio, kmm, gammastar, ns_star, ca, vpd, beta, gamma, cost_scalar, vcmax_start, gs_start, jmax_start ){
  out_optim <- optimr::optimr(
    par        = c( vcmax_start, gs_start, jmax_start ), # starting values
    lower      = c( vcmax_start*0.0001, gs_start*0.001, jmax_start*0.001 ),
    upper      = c( vcmax_start*20, gs_start*30, jmax_start*30 ),
    fn         = calc_optimal_gs_vcmax_jmax,
    args       = c(kmm, gammastar, ns_star, ca, vpd, beta, gamma),
    iabs       = (ppfd * fapar),
    kphio      = kphio,
    cost_scalar = cost_scalar,
    method     = "L-BFGS-B",
    maximize   = TRUE,
    control    = list( maxit = 1000 )
    )
  varlist <- calc_optimal_gs_vcmax_jmax( par=out_optim$par, args=c(kmm, gammastar, ns_star, ca, vpd, beta, gamma), iabs=(fapar*ppfd), kphio=kphio, cost_scalar=cost_scalar, maximize=FALSE, return_all=TRUE )
  return(varlist)
}

varlist_jmax <- wrap_calc_optimal_gs_vcmax_jmax( ppfd, fapar, kphio, out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta, gamma, unitcost_root$par, out_analytical$vcmax, out_analytical$gs, out_analytical$vcmax )

## Invoke function with optimised Vcmax and gs (in out_optim$par), now returning all variables
print( paste("Optimal chi from P-model:", out_analytical$chi ) )
print( paste("Optimal chi from Vcmax-gs optimization:", varlist$chi ) )
print( paste("Optimal chi from Vcmax-gs opt. with light limit.:", varlist_ll$chi ) )
print( paste("Optimal chi from Vcmax-gs opt. with Jmax limit.:", varlist_jmax$chi ) )
```

For a set of light levels
```{r}
ppfd_vec <- seq(from=0, to=8000, length.out=50)
out_opt_ll_list <- purrr::map(
  as.list( ppfd_vec),
  ~wrap_calc_optimal_gs_vcmax_ll( ., fapar, kphio, out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta, out_analytical$vcmax, out_analytical$gs )
  )
chi_vec_ll <- out_opt_ll_list %>% purrr::map_dbl("chi")

out_opt_jmax_list <- purrr::map(
  as.list( ppfd_vec),
  ~wrap_calc_optimal_gs_vcmax_jmax( ., fapar, kphio, out_analytical$kmm, out_analytical$gammastar, out_analytical$ns_star, out_analytical$ca, vpd, beta, gamma, unitcost_root$par, out_analytical$vcmax, out_analytical$gs, out_analytical$vcmax )
  )
chi_vec_jmax <- out_opt_jmax_list %>% purrr::map_dbl("chi")

plot( ppfd_vec, chi_vec_ll, type="l", xlab="PPFD", ylab="chi", ylim=c(0,1))
lines( ppfd_vec, chi_vec_jmax, col="blue" )
```
