---
title: "Usage"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

## Example P-model run

So much for the theory. Let's run the P-model, without $J_{\text{max}}$ limitation, for one set of inputs, being temperature, PPFD, VPD, CO$_2$, elevation, and fAPAR.

To do so, run the `rpmodel()` function from the rpmodel package:
```{r, message=FALSE, warning=FALSE}
library(rpmodel)
out_pmodel <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  elv            = 0,            # m.a.s.l.,
  kphio          = 0.05,         # quantum yield efficiency,
  beta           = 146,          # unit cost ratio a/b,
  fapar          = 1,            # fraction  ,
  ppfd           = 300,          # mol/m2/d,
  method_optci   = "prentice14",
  method_jmaxlim = "none",
  do_ftemp_kphio = FALSE 
  )
print( out_pmodel )

```

The function returns a list of variables (see also man page by `?rpmodel`), including $V_{\mathrm{cmax}}$, $g_s$, and all the parameters of the photosynthesis model ($K$, $\Gamma^{\ast}$), which are all internally consistent, as can be verified for...
$$
c_i = c_a - A / g_s = \chi c_a
$$

```{r}
print( out_analytical$ci )
print( out_analytical$ca - (out_analytical$gpp / c_molmass) / out_analytical$gs )
print( out_analytical$ca * out_analytical$chi )
```
Yes. 

And for...
$$
A = V_{\text{cmax}} \frac{c_i-\Gamma^{\ast}}{c_i + K} = \phi_0 I_{\text{abs}} \frac{c_i-\Gamma^{\ast}}{c_i + 2 \Gamma^{\ast}} = g_s (c_a - c_i)
$$

```{r}
print( out_analytical$gpp / c_molmass )
print( out_analytical$vcmax * (out_analytical$ci - out_analytical$gammastar) / (out_analytical$ci + out_analytical$kmm ))
print( out_analytical$gs * (out_analytical$ca - out_analytical$ci) )

print( kphio * ppfd * fapar * (out_analytical$ci - out_analytical$gammastar) / (out_analytical$ci + 2 * out_analytical$gammastar ))
```
Yes.